The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with \textit{Gaia} Data and Beyond

### 2019-05-10 paper/ms.tex
Update coordinate system discussion

\subsubsection{Galactic Orientation}
Second, one must define the angular orientation of the Galaxy. This was
defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
defining the coordinates of the Galactic center in B1950 coordinates as
(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
These two quantities, together with $R_0$, define the orientation of the
Galactic plane. However, there is growing evidence that the stellar midplane
is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
\citep{2019ApJ...871..145A}.  This tilt will contribute a linear dependence in Galactocentric radius to the systematic errors in actions induced by midplane offsets.

\subsubsection{Solar Height}
Third, one must define the Sun's vertical distance from the Galactic midplane,
which can be determined by identifying where the stellar density and
velocities reach a maximum (effectively the median height of all disk stars).
The Solar height is usually taken to be $\sim 25\,\pc$
\citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}. The discrepancy between gas-based and stellar-based determinations of the solar height is small, and thus only likely to dominate over intrinsic midplane variations on small scales, but will be relevant for detailed modeling of young and kinematically cold stars. For instance, it will induce a $\sim10\%$ error in $J_z$ for an
orbit with $z_{\text{max}}\sim100\,\pc$ (see Section~\ref{sec:ref_frame}).

\subsubsection{Local Standard of Rest}
Finally, one must define the LSR, or mean velocity of stars near the Sun
relative to the Galactic center (which is defined to have zero velocity), and
the velocity of the Sun relative to the LSR. The radial ($U_{\odot}$) and
vertical ($W_{\odot}$) components are computed by taking the mean motions of
different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
component ($V_{\odot}$) is more difficult to measure, but can be modeled using
the asymmetric drift relation \citep{2008gady.book.....B}. The values of the
components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. Their uncertainties should also lead to systematic errors in the actions, as given in Eqs. \ref{eq:induced_Jphi}--\ref{eq:induced_Jz}. For example, the
value of the circular velocity is taken to be $\sim 220\kms$
\citep[e.g.][]{2012ApJ...759..131B} with roughly 10\% uncertainty. Based on the analytic estimates in \S \ref{ssec:epi_action}, we expect this to translate to at least a 10\% systematic error in $J_{\phi}$.

>>>>

\subsubsection{Galactic Orientation}
Second, one must define the angular orientation of the Galaxy. This was
defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
defining the coordinates of the Galactic center in B1950 coordinates as
(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
These two quantities, together with $R_0$, define the orientation of the
Galactic plane. However, there is growing evidence that the stellar midplane
is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
\citep{2019ApJ...871..145A}.

This tilt will contribute a systematic offset in $z$, with the exact magnitude
depending on the position of the observed star. For instance,
\citet{2014ApJ...797...53G} quote a $\sim0.4\degree$ tilt at $3.1\,\kpc$,
corresponding to a vertical height of $\sim22\,\pc$. This corresponds to a
$37\%$ ($5\%$) error in $J_z$ for the thin-disk (thick-disk) orbit.

\subsubsection{Solar Height}
Third, one must define the Sun's vertical distance from the Galactic midplane,
which can be determined by identifying where the stellar density and
velocities reach a maximum (effectively the median height of all disk stars).
The Solar height is usually taken to be $\sim 25\,\pc$
\citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}. The discrepancy between gas-based and stellar-based determinations of the solar height is small, and thus only likely to dominate over intrinsic midplane variations on small scales, but will be relevant for detailed modeling of young and kinematically cold stars. For instance, it will induce a $\sim10\%$ error in $J_z$ for an
orbit with $z_{\text{max}}\sim100\,\pc$.

\subsubsection{Local Standard of Rest}
Finally, one must define the LSR, or mean velocity of stars near the Sun
relative to the Galactic center (which is defined to have zero velocity), and
the velocity of the Sun relative to the LSR. The radial ($U_{\odot}$) and
vertical ($W_{\odot}$) components are computed by taking the mean motions of
different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
component ($V_{\odot}$) is more difficult to measure, but can be modeled using
the asymmetric drift relation \citep{2008gady.book.....B}. The values of the
components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. Their uncertainties should also lead to systematic errors in the actions, as given in Eqs. \ref{eq:induced_Jphi}--\ref{eq:induced_Jz}. For example, the
value of the circular velocity is taken to be $\sim 220\kms$
\citep[e.g.][]{2012ApJ...759..131B} with roughly $10\%$ uncertainty. We expect this to translate to at least a $10\%$ systematic error in $J_{\phi}$.
